---
title: "Household Electric Power Consumption"
output: 
  html_notebook:
    highlight: tango
    theme: spacelab
    toc: yes
    toc_float: yes
    toc_depth: 4
---

```{r setup, include=FALSE }
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE)
```
##1 Frame the Problem


```{r message=FALSE, warning=FALSE}

library(caret)      #R modeling workhorse & ggplot2
library(tidyverse)  #Package for tidying datalibrary(magrittr)   #Enables piping
library(lubridate)  #Simplifies working with dates/times of a time series
library(VIM)        #Aids visualization and imputing of missing values
library(Hmisc)      #for descriptive statistics
library(GGally)     #ggcorr provides pretty cor plot
library(scales)
library(forecast)   #forcasting package

```
##2. Collect/Load Raw Data

```{r message=FALSE, cache=TRUE, results='hide'}

house_pwr <- read_delim('household_power_consumption.txt', col_names = TRUE, col_types = cols(Global_active_power='d', Global_reactive_power='d',
Voltage='d', Global_intensity='d', Sub_metering_1='d', Sub_metering_2='d', Sub_metering_3='d'), delim=';',  na='?')
```

```{r}
# Convert to a tibble
head(as_tibble(house_pwr))
```
##3 Process the Data

###3.1 Process Date and Time Features

```{r cache=TRUE}
house_pwr <- unite(house_pwr,Date, Time, col='DateTime', sep=' ')

house_pwr$DateTime <- as.POSIXct(house_pwr$DateTime, format="%d/%m/%Y %H:%M:%S", tz='America/New_York')
class(house_pwr$DateTime)
tz(house_pwr$DateTime)
```
###3.2 Rename Independent Variables
```{r}
colnames(house_pwr)[2] <- 'Glbl_actvPwr'
colnames(house_pwr)[3] <- 'Glbl_ractvPwr'
colnames(house_pwr)[6] <- 'Sub-Meter-1'
colnames(house_pwr)[7] <- 'Sub-Meter-2'
colnames(house_pwr)[8] <- 'Sub-Meter-3'

```
###3.3 Assess Missing Values
```{r cache=TRUE}
aggr_plot <- aggr(house_pwr, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(house_pwr),cex.axis=.7,
  gap=3, ylab=c("Histogram of missing data","Pattern"), digits=2)

housePWR_NA <- house_pwr[rowSums(is.na(house_pwr))>0,]
```
```{r}
# Remove rows with NA's
house_pwr <- na.omit(house_pwr)
sum(is.na(house_pwr))
```
###3.4 Create a Long Form of Data Set
```{r}
house_pwr_tidy <- house_pwr %>%
  gather(Meter, Watt_hr, `Sub-Meter-1`, `Sub-Meter-2`, `Sub-Meter-3`)  
```

```{r}
house_pwr_tidy$Meter <- factor(house_pwr_tidy$Meter)
glimpse(house_pwr_tidy)
```

##4 Explore the Data
###4.1 Visualizations of Energy Usage Across Sub-Meters and Time Periods
####4.1.1 Yearly Time Period
```{r cache=TRUE}
##-Year_Proportional Plot
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  group_by(year(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr)) %>%
  ggplot(aes(x=factor(`year(DateTime)`), sum, group=Meter,fill=Meter)) +
  labs(x='Year', y='Proportion of Energy Useage') +
  ggtitle('Proportion of Total Yearly Energy Consumption') +
  geom_bar(stat='identity', position='fill', color='black')
```
```{r cache=TRUE}
##-Year_Line Plot
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  group_by(year(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(`year(DateTime)`), sum, group=Meter, colour=Meter)) +
  labs(x='Year', y='kWh') +
  ggtitle('Total Yearly Energy Consumption') +
  geom_line(size=1)
```
```{r cached=TRUE}
##-Year Bar Chart
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  #mutate(year=year(DateTime)) %>%
  group_by(year(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr/1000),3)) %>%
  ggplot(aes(x=factor(`year(DateTime)`), y=sum)) +
  labs(x='Year', y='kWh') +
  ggtitle('Total Energy Useage by Year') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black')
```
####4.1.2 Quarterly Time Period


```{r cached = TRUE}
#Quarter bar  plot
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  #filter(year(DateTime)<2010) %>%
  #mutate(quarter=quarter(DateTime)) %>%
  group_by(quarter(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr/1000),3)) %>%
  ggplot(aes(x=factor(`quarter(DateTime)`), y=sum)) +
  labs(x='Quarter of the Year', y='Wh') +
  ggtitle('Total Quarterly Energy Consumption') +
  geom_bar(stat='identity', aes(fill = Meter), color='black')
```

####4.1.3 Monthly Time Period


```{r cached=TRUE}
###-Month- Line Plot
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  mutate(Month=lubridate::month(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Month, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Month), sum, group=Meter, colour=Meter)) +
  labs(x='Month of the Year', y='Wh') +
  ggtitle('Average Monthly Energy Useage') +
  geom_line(size=1) +
  geom_line()
```

```{r echo=TRUE, cached=TRUE}
###-Month bar chart
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  #filter(year(DateTime)<2010) %>%
  mutate(Month=lubridate::month(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Month, Meter) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3) %>%
  ggplot(aes(x=factor(Month), y=sum)) +
  labs(x='Month of the Year', y='kWh') +
  ggtitle('Total Energy Useage by Month of the Year') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black')
```
####4.1.4 Week of the Year Time Period

```{r cached=TRUE}
#Week of the year- line plot
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  group_by(week(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3) %>%
  ggplot(aes(x=factor(`week(DateTime)`), sum, group=Meter,colour=Meter)) +
  labs(x='Week of the Year', y='kWh') +
  ggtitle('Total Energy Usage by Week of the Year') +
  theme(axis.text.x = element_text(angle=90)) +
  geom_line(size=1) +
  geom_line()
```
```{r cached=TRUE}
#Week of the year- bar plot
house_pwr_tidy %>%
  #filter(year(DateTime)>2006) %>%
  group_by(week(DateTime), Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(`week(DateTime)`), y=sum)) +
  labs(x='Week of the Year', y='kWh') +
  ggtitle('Total Energy Usage by Week of the Year') +
  theme(axis.text.x = element_text(angle=90)) +
  geom_bar(stat='identity', aes(fill=Meter), colour='black')

```
####4.1.5 Day of the Week Time Period
```{r cached = TRUE}
###-Day of Week- Porportional Plot
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Day, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Day), sum, group=Meter,fill=Meter)) +
  labs(x='Day of the Week', y='Proportion of Energy Useage') +
  ggtitle('Proportion of Total Daily Energy Consumption') +
  geom_bar(stat='identity', position='fill', color='black')
```

```{r cached=TRUE}
###-Day of Week- Line Plot
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Day, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Day), sum, group=Meter, colour=Meter)) +
  labs(x='Day of the Week', y='Wh') +
  ggtitle('Average Daily Energy Consumption') +
  geom_line(size=1) +
  geom_line()

```

```{r cached=TRUE}
###-Day of Week bar chart
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  mutate(Day=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(Day, Meter) %>%
  summarise(sum=sum(Watt_hr/1000)) %>%
  ggplot(aes(x=factor(Day), y=sum)) +
  labs(x='Day of the Week', y='kWh') +
  ggtitle('Total Energy Useage by Day of Week') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black')

```
####4.1.7 Hour of the Day Time Period

```{r cached=TRUE}
###-Hour of the Day-Line Plot
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  group_by(hour(DateTime), Meter) %>%
  summarise(sum=round(mean(Watt_hr), 3)) %>%
  ggplot(aes(x=factor(`hour(DateTime)`), sum, group=Meter,colour=Meter)) +
  labs(x='Hour of the Day', y='Wh') +
  ggtitle('Average Hourly Energy Consumption') +
  geom_line(size=1) +
  geom_line()
```

```{r cached=TRUE}
#Hour of day bar chart
house_pwr_tidy %>%
  filter(year(DateTime)>2006) %>%
  group_by(hour(DateTime), Meter) %>%
  summarise(sum=round(sum(Watt_hr)/1000),3) %>%
  ggplot(aes(x=factor(`hour(DateTime)`), y=sum)) +
  labs(x='Hour of the Day', y='kWh') +
  ggtitle('Total Energy Useage by Hour of the Day') +
  geom_bar(stat='identity', aes(fill = Meter), colour='black')
```
##5 Subset Data Set by Time Periods of Interest
###5.1 Quarterly 
Subset by Quarter
```{r}
housePWR_qtr <- house_pwr %>%
  filter(year(DateTime) > 2006) %>%
  #filter(year(DateTime)<2010) %>%
  group_by(year(DateTime), quarter(DateTime)) %>%
  summarise(Sub_Meter_1=round(sum(`Sub-Meter-1`/1000, na.rm=TRUE), 3),
            Sub_Meter_2=round(sum(`Sub-Meter-2`/1000, na.rm=TRUE), 3),
            Sub_Meter_3=round(sum(`Sub-Meter-3`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))
```
###5.2 Monthly 
```{r}
housePWR_mnth <- house_pwr %>%
  #filter(year(DateTime) > 2006) %>%
  #filter(year(DateTime)<2011) %>%
  group_by(year(DateTime), month(DateTime)) %>%
  summarise(Sub_Meter_1=round(sum(`Sub-Meter-1`/1000, na.rm=TRUE), 3),
            Sub_Meter_2=round(sum(`Sub-Meter-2`/1000, na.rm=TRUE), 3),
            Sub_Meter_3=round(sum(`Sub-Meter-3`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))
```
###5.3 Weekly
```{r}
housePWR_wkofYr <- house_pwr %>%
  #filter(year(DateTime)>2006) %>%
  #filter(year(DateTime)<2010) %>%
  mutate(DofWk=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(year(DateTime),week(DateTime)) %>%
  summarise(Sub_Meter_1=round(sum(`Sub-Meter-1`/1000, na.rm=TRUE), 3),
            Sub_Meter_2=round(sum(`Sub-Meter-2`/1000, na.rm=TRUE), 3),
            Sub_Meter_3=round(sum(`Sub-Meter-3`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))

```
###5.4 Day of Week
```{r}
housePWR_dofWk <- house_pwr %>%
  filter(year(DateTime)>2006) %>%
  #filter(year(DateTime)<2010) %>%
  mutate(DofWk=lubridate::wday(DateTime, label=TRUE, abbr=TRUE)) %>%
  group_by(week(DateTime),DofWk) %>%
  summarise(Sub_Meter_1=round(sum(`Sub-Meter-1`/1000, na.rm=TRUE), 3),
            Sub_Meter_2=round(sum(`Sub-Meter-2`/1000, na.rm=TRUE), 3),
            Sub_Meter_3=round(sum(`Sub-Meter-3`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))
```
###5.5 Hour of the Day
```{r}
housePWR_hofDay <- house_pwr %>%
  filter(year(DateTime)>2006) %>%
  #filter((minute(DateTime) %% 5) == 0) %>%
  group_by(wday(DateTime), hour(DateTime)) %>%
  summarise(Sub_Meter_1=round(sum(`Sub-Meter-1`/1000, na.rm=TRUE), 3),
            Sub_Meter_2=round(sum(`Sub-Meter-2`/1000, na.rm=TRUE), 3),
            Sub_Meter_3=round(sum(`Sub-Meter-3`/1000, na.rm=TRUE), 3),
            first_DateTime = first(DateTime))

```
##6 Convert to Time Series and Plot TS
###6.1 Quarterly Time Series
```{r}
housePWR_qtrTS <- ts(housePWR_qtr[,3:5], frequency=4, start=c(2007,1), end=c(2010,3))
plot(housePWR_qtrTS, plot.type='s',
     #xaxt='n',
     #xlim=c(2007, 2010),
     #xaxp=c(2006, 2011, 5),
     col=c('red', 'green', 'blue'),
     main='Total Quarterly kWh Consumption',
     xlab='Year', ylab = 'kWh')
minor.tick(nx=4)
b <- c('Sub-meter-1', 'Sub-meter-2', 'Sub-meter-3')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')
```
###6.2 Monthly Time Series
```{r}
housePWR_mnthTS <- ts(housePWR_mnth[,3:5], frequency = 12, start=c(2007,1), end=c(2010,11))
plot(housePWR_mnthTS, plot.type='s',#xaxt='n',
     #xaxp = c(1,13,12),
     col=c('red', 'green', 'blue'),
     main='Total Monthly kWh Consumption',
     xlab='Year/Month', ylab = 'kWh')
minor.tick(nx=12)
b <- c('Sub-meter-1', 'Sub-meter-2', 'Sub-meter-3')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')
```
###6.3 Week of the Year Time Series
```{r}
housePWR_wkofYrTS <- ts(housePWR_wkofYr[,3:5], frequency =53, start=c(2006,51), end=c(2010,47))
plot(housePWR_wkofYrTS, plot.type='s', #xaxt='n',
     #xaxp = c(1, 13, 12),
     col=c('red', 'green', 'blue'),
     xlab ='Year', ylab = 'kWh',
     #ylim=c(0,120),
     main='Total Energy Consumption by Week of the Year')
#axis(side=1, at= c(1, 2,3,4,5,6,7,8,9,10,11,12,13), labels=MonthLst)
minor.tick(nx=52)
b <- c('Sub-meter-1', 'Sub-meter-2', 'Sub-meter-3')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')
```
###6.4 Day of the Week Time Series
```{r}
housePWR_hofDayTS <- ts(housePWR_hofDay[,3:5], frequency=24, start=c(0,0))
plot(housePWR_hofDayTS, plot.type='s',
     #xaxp = c(0,48, 47),
     col=c('red', 'green', 'blue'),
     xlab='Day', ylab = 'kWh',
     main='Total kWh Consumption by Hour of the Day')
minor.tick(nx=24)
b <- c('Sub-meter-1', 'Sub-meter-2', 'Sub-meter-3')
legend('topleft', b, col=c('red', 'green', 'blue'), lwd=2, bty='n')
```
##7 Forecasts of Energy Consumption_Linear Model
###7.1 Quarterly Forecast
```{r}
fit1 <- tslm(housePWR_qtrTS[,3] ~ trend + season)
x <- forecast(fit1, h=4, level=c(80,95))
plot(x, showgap=FALSE, include=3,
     shadecols=c('slategray3','slategray'),
     xlab='Year', ylab='kWh',
     main='4-Quarter Forecast of Quartlerly Energy Consumption for Submeter-3')
minor.tick(nx=2)
summary(x)
```
```{r}
plot.ts(x=fit1$fitted.values, y=housePWR_qtrTS[,3], xy.lines = FALSE,
        xy.labels = FALSE,
        xlab='Fitted Value',
        ylab='Actual',
        main='Quarterly Predicted vs. Actual Values')
abline(0,1, col='blue')

```

```{r}
plot.ts(x=fit1$fitted.values, y=fit1$residuals, xy.lines=FALSE,
        xy.labels = FALSE,
        xlab='Predicted Value',
        ylab='Residuals',
        main='Quarterly Predicted vs. Residuals')
abline(0,0, lty=2, col='grey')
```

```{r}
qqnorm(fit1$residuals, main='Quarterly Q-Q Plot')
qqline(fit1$residuals)
```
 
```{r}
checkresiduals(fit1)

```
###7.2 Monthly Forecast
```{r}
fit2 <- tslm(housePWR_mnthTS[,3] ~ trend + season)
y <- forecast(fit2,h=12, level=c(80,95))
plot(y, showgap=FALSE, include=5,
  shadecols=c('slategray3','slategray'),
  xlab ='Year',
  ylab=' kWh',
  main='12-Month Forecast of Monthly Energy Consumption')
minor.tick(nx=6)
```
```{r}
plot.ts(x=fit2$fitted.values, y=housePWR_mnthTS[,3], xy.lines = FALSE,
        xy.labels = FALSE,
        xlab='Fitted Value',
        ylab='Actual',
        main='Monthly Predicted vs. Actual Values')
abline(0,1, col='blue')
```

```{r}
plot.ts(x=fit2$fitted.values, y=fit2$residuals, xy.lines=FALSE,
        xy.labels = FALSE,
        xlab='Predicted Value',
        ylab='Residuals',
        main='Monthly Predicted vs. Residuals')
abline(0,0, lty=2, col='grey')
```
```{r}
qqnorm(fit2$residuals, main='Monthly Q-Q Plot')
qqline(fit2$residuals)

```
```{r}
checkresiduals(fit2)

```
```{r}
fit3 <- tslm(housePWR_wkofYrTS[,3] ~ trend + season)
z <- forecast(fit3, level=c(80,95), h=24)
plot(z, showgap=FALSE, include=8,
     shadecols=c('slategray3','slategray'),
     xlab ='Year',
     ylab='kWh',
     main='24-Week Forecast of Weekly Energy Consumption on Submeter-3')
```
```{r}
plot.ts(x=fit3$fitted.values, y=housePWR_wkofYrTS[,3], xy.lines = FALSE,
        xy.labels = FALSE,
        xlab='Fitted Value',
        ylab='Actual',
        main='Week of Year Predicted vs. Actual Values')
abline(0,1, col='blue')
```
```{r}
plot.ts(x=fit3$fitted.values, y=fit3$residuals, xy.lines=FALSE,
        xy.labels = FALSE,
        xlab='Predicted Values',
        ylab='Residuals',
        main='Week of Year Predicted vs. Residuals')
abline(0,0, lty=2, col='grey')
```
```{r}
qqnorm(fit3$residuals, main='Weekly Q-Q Plot')
qqline(fit3$residuals)
```
```{r}
checkresiduals(fit3, main='Residuals from Week of Year Linear Regression Model')

```
##8 Holt-Winters Simple Exponential Smoothing Model
###8.1 Quarterly Time Series
```{r}
##-Sub-Meter-3
#-Decompose TS
qtr_decomp3 <- decompose(housePWR_qtrTS[,3])
autoplot(qtr_decomp3,  range.bars = TRUE) +
  xlab('Year') +
  ylab('kWh') +
  ggtitle('Decomposed Quarterly Time Series- Sub-Meter-3')
```
```{r}
#-remove seasonality
qtr_seasonAdj3 <- seasadj(qtr_decomp3)

plot(qtr_seasonAdj3, #xaxt='n',
     col='blue',
     xlab='Year', ylab='kWh',
     #xlim=c(2007, 2010),
     xaxp=c(2006, 2010, 4),
     main='Seasonally Adjusted Quarterly Time Series')
minor.tick(nx=2)
b <-'Sub-meter-3'
legend('topleft', b, col='blue', lwd=2, bty='n')
```

```{r}
#-Fit Holt Winters simple exponetial smoothing model
qtr_smooth3 <- HoltWinters(qtr_seasonAdj3,
                           beta=FALSE,
                           gamma=FALSE)
plot(qtr_smooth3, col='blue', #xaxt='n',
     xlab='Year', ylab = 'kWh',
     xlim=c(2007, 2011),
     xaxp=c(2006, 2010, 4),
     main='Simple Exponential Smoothing Holt-Winters Model for Quarterly Time Series')
minor.tick(nx=4)
legend('topleft', 'Sub-Meter-3', col='blue', lwd=2, bty='n')
```

```{r}
#-Forecast
qtr_smoothFcast3 <- forecast(qtr_smooth3, h=5,level = c(80,95))

#summary(qtr_smoothFcast3)
checkresiduals(qtr_smoothFcast3)
plot(qtr_smoothFcast3, include=1,
     #xaxt='n',
     shadecols=c('slategray3','slategray'),
     col='blue',
     #xaxp=c(2010.6,2011.5,4),
     xlab='Year', ylab = 'Wh',
     #xlim=c(2010,2011),
     main='5 Quarter Forecast of Energy Useage on Sub-Meter 3')
#minor.tick(nx=4)
#axis(side=1, at= c(1, 2,3,4,5,6,7,8,9,10,11,12, 13), labels=MonthLst)
legend('topleft', 'Sub-Meter-3', col='blue', lwd=2, bty='n')
```





