---
title: "Analysis of Submetered Household Electric Power Usage"
author: "EJG"
date: "January 2, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(caret)  
library(tidyverse)  #Package for tidying data
library(kableExtra) #fancy table generator
library(knitr)      #report generation
```


## Abstract

This project is analyzed from the perspective of a data scientist working for an "Internet of Things" analytics consulting firm whose client is a home developer.  The client would like to know if insights can be found in the sub-metered energy consumption data set that could be used as an incentive to potential home buyers that may be interested in "smart home" technology.

Turning the business problem into a data science problem, the goal was to find examples of the three deliverables below to support the business objective.


I.  *Sub-metered energy consumption data that provides enough granularity to uncover trends in behavior or appliance performance.*  

II. *Identification of peak energy usage can be identified allowing for the potential to modify behavior to take advantage of off-peak electricity rates.*

III. *Longer-term patterns of energy usage that can be used to predict future usage with the potential to flag appliance degradation or unusual energy consumption.*


## Methods
The 'household_power_consumption' data set is available at the [UCI Machine Learning Repository](http://archive.ics.uci.edu/ml/datasets/Individual+household+electric+power+consumption).  It consists of time series data with 2,075,259 energy measurements and 9 variables.  The variable definitions are shown below.

```{r echo=FALSE, message=FALSE}
#-Read in CSV file of data feature definitions
def_table <- read.csv('Energy_submeter_defs.csv')
  
#-create table of feature definitions
kable(def_table, 
      col.names=c('Feature', 'Definition', 'Sub-Meter-Coverage'),       caption='Data Set Feature Definitions', 'html') %>% 
  kable_styling(bootstrap_options = c('striped', 'hover', 'responsive', 'bordered'), 
                full_width = F,
                font_size = 14)
```

